### 2020-09-14 paper/coloringBergeGraphs.tex
Plots and appendix
Progress

We will now show how to find a stable set in $G$ of size $\alpha(G)$.

@@

  \noindent Output: A maximum cardinality stable set in $G$.
\end{alg}
\begin{algtext2}
  Let $v_1, \ldots v_n$ be an ordering of vertices of $G$. We will construct a sequence of induced subgraphs $G = G_0 \supseteq G_1 \supseteq \ldots \supseteq G_n$, so that $G_n$ is a required stable set.

  Let $G_0 \leftarrow G$. Then, for each $i \geq 1$, compute $\alpha(G_{i-1} \setminus v_i)$. If $\alpha(G_{i-1} \setminus v_i) = \alpha(G)$, then set $G_i \leftarrow G_{i-1} \setminus \{v_i\}$, else set $G_i \leftarrow G_{i-1}$.

  Return $G_n$.
\end{algtext2}

Let us prove that $G_n$ is indeed a stable set. Suppose otherwise and let $v_iv_j$ be an edge in $G_n$ with $i < j$ and $i$ minimal. But then $\alpha(G_{i-1} \setminus v_i) = \alpha(G_{i-1}) = \alpha(G)$ so by our construction $v_i$ is not in $G_i$ and $v_iv_j$ is not an edge of $G_n$. Therefore there are no edges in $G_n$.

Because at every step we have $\alpha(G_i) = \alpha(G_{i-1})$, therefore $\alpha(G_n) = \alpha(G)$, so $G_n$ is required maximum cardinality stable set.

The running time of \cref{alg:maxStableSet} is polynomial, because we construct $n$ auxiliary graphs, each requiring calculating $\alpha$ once plus additional $O(|V|^2)$ time for constructing the graph.

Given a weight function $w : V \rightarrow \mathbb{N}$ we could calculate the maximum weighted stable set in $G$ in the following manner. Create graph $G'$ by replacing every node $v$ by a set $W_v$ of $w(v)$ nonadjacent nodes, making two nodes $x \in W_v$, $y \in W_u$ adjacent in $G'$ if and only if the nodes $v$, $u$ are adjacent in $G$. Then calculate a maximum cardinality stable set in $G'$ (we remark that $G'$ is still perfect because every new introduced hole is even) and return a result of those vertices in $G$ whose any (and therefore all) copies were chosen. We will use this technique later on.

\subsection{Stable set intersecting all maximum cardinality cliques}
Next, let us show how to find a stable set intersecting all the maximum cardinality cliques of $G$.
\begin{alg}
  \label{alg:ssIntersectingCliques}
  Input: A perfect graph $G = (V, E)$.

  \noindent Output: A stable set which intersects all the maximum cardinality cliques of $G$.
\end{alg}
\begin{algtext2}
  We will create a list $Q_1, \ldots Q_t$ of all maximum cardinality cliques of $G$.

  Let $Q_1 \leftarrow$ a maximum cardinality clique of $G$. We calculate this by running \cref{alg:maxStableSet} on $\overline{G}$.

  Now suppose $Q_1, \ldots, Q_t$ have been found. We show how to calculate $Q_{t+1}$ or see that we are done.

  Let us define a weight function $w : V \rightarrow \mathbb{N}$, so that for $v \in V$, $w(v)$ is equal to the number of cliques $Q_1, \ldots Q_t$ that contain $v$.

  Assign $S \leftarrow$ the maximum $w$-weighted stable set, as described in a remark to \cref{alg:maxStableSet}. It is easy to see that $S$ has weight $t$, which means that $S$ meets each of $Q_1, \ldots Q_t$.

  If $\omega(G \setminus S) < \omega(G)$, then $S$ meets all the maximum cardinality cliques in G so we return $S$. Otherwise we find a maximum cardinality clique in $G \setminus S$ (it will be of size $\omega(G)$, because $\omega(G \setminus S) = \omega(G)$), add it to our list as $Q_{t+1}$ and continue with longer list.

\end{algtext2}

@@

  \noindent Output: A coloring of $G$ using $\chi(G)$ colors.
\end{alg}
\begin{algtext2}
  If $G$ is equal to its maximum cardinality stable set, color all vertices one color and return.

  Else find $S$ intersecting all maximum cardinality cliques of $G$ (\cref{alg:ssIntersectingCliques}). Color recursively all vertices of $G \setminus S$ with $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$ colors and all vertices of $S$ with one additional color.
\end{algtext2}

We will call recursion at most $O(|V|)$ times, each step of recursion is polynomial in time. Therefore the running time of \cref{alg:minColoring} is polynomial.

@@

A \emph{prism} is a graph consisting of two disjoint triangles and two disjoint paths between them. Notice, that for a graph to contain no odd hole, all three paths in a prism must have the same parity. A prism with all three paths odd is called an \emph{odd prism}.

>>>>

We will now show how to find a stable set in $G$ of size $\alpha(G)$. We will construct a sequence of induced subgraphs $G = G_0 \supseteq G_1 \supseteq \ldots \supseteq G_{|V|}$, so that $G_{|V|}$ is a required stable set.

@@

  \noindent Output: A maximum cardinality stable set in $G$.
\end{alg}

\begin{algorithmic}[1]
  \mProcedure{Get-Max-Stable-Set}{$G$}
  \ls $v_1, \ldots, v_n \gets$ an ordering of $V(G)$
  \ls $G_0 \leftarrow G$
  \ls $\alpha_G = \alpha(G)$
  \mForEach{$1 \leq i \leq |V|$}
    \mIf{$\alpha(G_{i-1} \setminus v_i) = \alpha_G$}
      \ls $G_i \leftarrow G_{i-1} \setminus \{v_i\}$
    \mElse
      \ls $G_i \leftarrow G_{i-1}$
    \mEndIf
  \mEndFor
  \ls \RETURN $G_{|V|}$
  \mEndProcedure
\end{algorithmic}

\begin{theorem}
  \Cref{alg:maxStableSet} works in polynomial time and for an input of $G$ returns a maximal cardinality stable set of $G$.
\end{theorem}
\begin{proof}
  Let us prove that $G_{|V|}$ is indeed a stable set. Suppose otherwise and let $v_iv_j$ be an edge in $G_{|V|}$ with $i < j$ and $i$ minimal. But then $\alpha(G_{i-1} \setminus v_i) = \alpha(G_{i-1}) = \alpha(G)$ so by our construction $v_i$ is not in $G_i$ and $v_iv_j$ is not an edge of $G_n$. Therefore there are no edges in $G_{|V|}$.

  Because at every step we have $\alpha(G_i) = \alpha(G_{i-1})$, therefore $\alpha(G_{|V|}) = \alpha(G)$, so $G_{|V|}$ is required maximum cardinality stable set.

  The running time of \cref{alg:maxStableSet} is polynomial, because we construct $|V|$ auxiliary graphs, each requiring calculating $\alpha$ once plus additional $O(|V|^2)$ time for constructing the graph.
\end{proof}

Given a natural weight function $w : V \rightarrow \mathbb{N}$ we could calculate the maximum weighted stable set in $G$ using the following modification:

\begin{alg}[maximum weighted stable set in a perfect graph]
  \label{alg:maxWeightedStableSet}
  Input: A perfect graph $G = (V, E)$ and a weight function $w : V \rightarrow \mathbb{N}$.

  \noindent Output: A maximum weighted stable set in $G$.
\end{alg}

\begin{algorithmic}[1]
  \mProcedure{Get-Max-Weighted-Stable-Set}{$G, w$}
  \mForEach{$v \in V(G)$}
    \ls $W_v \gets$ a set of nonadjacent vertices of size $w(v)$
  \mEndFor
  \ls $V' \gets \bigcup W_v$, for $v \in V(G)$ \Comment{$|V'| = \sum_{v\in V(G)} w(v)$ \vspace{-.1cm}}
  \ls $E' \gets \{xy$ for $x, y \in V'$, such that for some $u, v \in V, u \neq v$:
  \lsx~~~~~~~~ $x \in W_u$, $y \in W_v$, and $uv \in E(G)\}$
  \ls $G' \gets (V', E')$
  \ls $S' \gets$ \textsc{Get-Max-Stable-Set$(G')$}
  \ls $S \gets \{v \in V(G)$, such that for some $x\in W_v$, $x \in S'\}$
  \ls \RETURN $S$
  \mEndProcedure
\end{algorithmic}

We create graph $G'$ by replacing every node $v$ by a set $W_v$ of $w(v)$ nonadjacent nodes, making two nodes $x \in W_v$, $y \in W_u$ adjacent in $G'$ if and only if the nodes $v$, $u$ are adjacent in $G$. Then we can calculate a maximum cardinality stable set in $G'$ (we remark that $G'$ is still perfect because every new introduced hole is even) and return a result of those vertices in $G$ whose any (and therefore all) copies were chosen.

\subsection{Stable set intersecting all maximum cardinality cliques}
Next, let us show how to find a stable set intersecting all the maximum cardinality cliques of $G$. We will create a list $Q_1, \ldots Q_t$ of all maximum cardinality cliques of $G$ and return a stable set intersecting them all.

Recall that to calculate a maximum cardinality clique of $G$, we can calculate a maximum cardinality stable set of $\overline{G}$.
\begin{alg}
  \label{alg:ssIntersectingCliques}
  Input: A perfect graph $G = (V, E)$.

  \noindent Output: A stable set which intersects all the maximum cardinality cliques of $G$.
\end{alg}

\begin{algorithmic}[1]
  \mProcedure{Get-Stable-Set-Intersecting-Max-Cliques}{$G$}
    \ls $\omega_G \gets \omega(G)$
    \ls $Q_1 \gets$ \textsc{Max-Stable-Set}$(\overline{G})$
    \ls $t \gets 1$
    \mWhile{true}
      \ls $w \gets$ a weight function $V \rightarrow \mathbb{N}$, so that for $v \in V$, $w(v)$ is equal to
      \lsx the number of cliques $Q_1, \ldots Q_t$ that contain $v$.
      \ls $S \gets$ \textsc{Max-Weighted-Stable-Set}$(G, w)$ \label{line:ssGetS}
      \mIf{$\omega(G \setminus S) < \omega_G$}
        \ls \RETURN $S$
      \mElse
        \ls $t \gets t+1$ \vspace{-.1cm}
        \ls $Q_t \gets$ \textsc{Max-Stable-Set}$(\overline{G \setminus S})$
      \mEndIf
    \mEndWhile
  \mEndProcedure
\end{algorithmic}

To understand \Cref{alg:ssIntersectingCliques}, let us note that $S$ calculated in line \ref{line:ssGetS} has a weight equal $t$, which means that $S$ meets each of $Q_1, \ldots Q_t$. Then, if $\omega(G \setminus S) < \omega(G)$, then $S$ meets all the maximum cardinality cliques in G so we return $S$.

@@

  \noindent Output: $C$ -- a coloring of $G$ using $\chi(G)$ colors.
\end{alg}

\begin{algorithmic}[1]
  \mProcedure{Color-Perfect-Graph}{$G$}
  \mIf{$G =$~\textsc{Get-Max-Stable-Set$(G)$}}
    \ls $C \gets$ coloring all vertices of $G$ with one color
    \ls \RETURN $C$
  \mElse
    \ls $S \gets$ \textsc{Get-Stable-Set-Intersecting-Max-Cliques}$(G)$
    \ls $C' \gets$ \textsc{Color-Perfect-Graph}$(G \setminus S)$
    \ls $C \gets C'$ with vertices of $S$ colored with an additional color
    \ls \RETURN $C$
  \mEndIf
  \mEndProcedure
\end{algorithmic}

Note that coloring $C'$ uses $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$. We will call recursion at most $O(|V|)$ times, each step of recursion is polynomial in time. Therefore the running time of \cref{alg:minColoring} is polynomial.

@@

\begin{defn}[prism]
  A \emph{prism} is a graph consisting of two disjoint triangles and two disjoint paths between them. Notice, that for a graph to contain no odd hole, all three paths in a prism must have the same parity. A prism with all three paths odd is called an \emph{odd prism}.
  \label{def:prism}
\end{defn}

### 2020-09-14 paper/experiments.tex
Plot
save
Progress

\begin{figure}
  \centering
  \input{plots/biparite.t.in.out.csv.lines.pgf}
  \caption{TODO: redo with more data Random bipartite graphs}
  \label{plot:perf2Lines}
\end{figure}

@@

\begin{figure}
  % \begin{subfigure}{.5\textwidth}
  %   \centering
  %   \input{plots/color.perf.t.in.out.csv.lines.pgf}
  %   \caption{Random perfect graphs}
  % \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Random perfect graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Line graphs of random perfect graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.grid5by4to9.t.in.out.csv.lines.pgf}
    \caption{Lattice graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.rookGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Rook graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.knightGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Knight graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering%
    \input{plots/color.hypercube20to40.t.in.out.csv.lines.pgf}
    \caption{Hypercube graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Split graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Full binary trees}
  \end{subfigure}
\end{figure}

>>>>

\begin{figure}
  \centering
  \input{plots/biparite.t.in.out.csv.lines.pgf}
  \caption{Random bipartite graphs}
  \label{plot:perf2Lines}
\end{figure}

@@

\begin{figure}
  % \begin{subfigure}{.5\textwidth}
  %   \centering
  %   \input{plots/color.perf.t.in.out.csv.lines.pgf}
  %   \caption{Random perfect graphs}
  % \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.biparite18to48.t.in.out.csv.lines.pgf}
    \caption{Random bipartite graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.perfLin.t.in.out.csv.lines.pgf}
    \caption{Line graphs of random bipartite graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.grid5by4to9.t.in.out.csv.lines.pgf}
    \caption{Lattice graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.rookGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Rook graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.knightGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Knight graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering%
    \input{plots/color.hypercube20to40.t.in.out.csv.lines.pgf}
    \caption{Hypercube graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.split.t.in.out.csv.lines.pgf}
    \caption{Split graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.fullBinary20to45.t.in.out.csv.lines.pgf}
    \caption{Full binary trees}
  \end{subfigure}
  \caption{Overall times of CSDP Color}
\end{figure}
